Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-31 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: more edits to dcs section

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{equation} \label{eq:action_dcs}
    S_{\rm CS} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\nabla \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm CS} \, \vartheta \, {}^{*}RR
    \right],
\end{equation}
%
where $\vartheta$ is a pseudoscalar field, which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R^{\mu}{}_{\nu}{}^{\rho\sigma} R^{\nu}_{\mu\rho\sigma}$,
%
where ${}^{*}R^{\mu}{}_{\nu}{}^{\rho\sigma}$ is the dual of the Riemann tensor,
defined as
%
${}^{*}R^{\mu}{}_{\nu}{}^{\rho\sigma} =
\epsilon^{\mu}{}_{\nu\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and where $\epsilon^{\mu}{}_{\nu\gamma\delta}$ is the Levi-Civita tensor.
%
As in scalar-Gauss-Bonnet theories, $\int \dV {}^{*}RR$ alone is a boundary
term~\cite{Jackiw:2003pm}, thus does not contribute to the field equations in
four-dimensions.
%
However, this terms can modify the theory's field equation by coupling it to
$\vartheta$.
%
The coupling strength set by the coupling constant $\ell_{\rm CS}$, with
dimensions of length.

The theory admits same the garden-variety Schwarzschild BH solution of
GR~\cite{Jackiw:2003pm}, but admits non-Kerr solutions for spinning BHs~\cite{Yunes:2009hc,Konno:2009kg}.
%
These BHs support a scalar field which asymptotically falls-off as $r^{-2}$, to which we can associate
a scalar dipole charge.
%
As a consequence, the leading-scalar radiative channel of BBH to lose energy is quadrupolar,
which affects the GW phase at 2PN order~\cite{Yagi:2011xp,Alexander:2017jmt}.
%
Deviations from GR at this PN order are poorly constrained with present GW
observations~\cite{LIGOScientific:2021sio}, making dCS gravity unconstrained by
analyses that rely on the inspiral portion of GWs from BBH
binaries~\cite{Nair:2019iur,Perkins:2021mhb}.
%
However, the theory has been constrained by Ref.~\cite{Silva:2020acr}, that found
%
$\ell_{\rm CS} \leqslant 8.5$~km,
%
by folding data from the x-ray observations of the isolated neutron star (NS)
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
neutron star (BNS) GW170817~\cite{TheLIGOScientific:2017qsa}.

>>>>

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{equation} \label{eq:action_dcs}
    S_{\rm CS} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\nabla \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm CS} \, \vartheta \, {}^{*}RR
    \right],
\end{equation}
%
where $\vartheta$ is a pseudoscalar field, which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor,
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and where $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
As in scalar-Gauss-Bonnet theories, $\int \dV {}^{*}RR$ alone is a boundary
term~\cite{Jackiw:2003pm}, thus does not contribute to the field equations in
four-dimensions.
%
However, this terms can modify the theory's field equation by coupling it to
$\vartheta$.
%
The coupling strength is set by the coupling constant $\ell_{\rm CS}$, with
dimensions of length.

The theory admits as solution the garden-variety Schwarzschild BH of GR. This
is not case when rotation is included and the Kerr metric is not a solution of
the theory~\cite{Jackiw:2003pm}.
%
These rotating BHs support a scalar field which falls-off as
$r^{-2}$ asymptically, to which we can associate a scalar dipole
charge~\cite{Yunes:2009hc,Konno:2009kg}.
%
As a consequence, the leading-scalar radiative channel of BBH to lose energy is quadrupolar,
which affects the GW phase at 2PN order~\cite{Yagi:2011xp,Alexander:2017jmt}.
%
Deviations from GR at this PN order are poorly constrained with present GW
observations~\cite{LIGOScientific:2021sio}, making dCS gravity unconstrained by
analyses that rely on the inspiral portion of GWs from BBH
binaries~\cite{Nair:2019iur,Perkins:2021mhb}.
%
Nonetheless, the theory has been constrained by Ref.~\cite{Silva:2020acr}, that found
%
$\ell_{\rm CS} \leqslant 8.5$~km,
%
by folding data from the x-ray observations of the isolated neutron star (NS)
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
neutron star (BNS) GW170817~\cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki}.
